### 2017-04-02 Appendix/appendix.tex
Added the Appendix

\documentclass[a4paper,11pt]{article}
\usepackage{amstext}
\usepackage{amsmath}

\begin{document}

\appendix
\section{Recursive computations of the ``NMix'' likelihood}

The likelihood for the simplest case is
\begin{displaymath}
    \text{Prob}(y) = \sum_{n = y}^{\infty}
    \text{Pois}(n ; \lambda) \;\times\; \text{Binom}(y;  n, p)
\end{displaymath}
where $\text{Poiss}(n; \lambda)$ is the density for the Poisson
distribution with mean $\lambda$, $\lambda^{n}\exp(-\lambda)/n!$, and
$\text{Binom}(y; n, p)$ is the density for the Binomial distribution
with $n$ trials and probability $p$, ${n \choose y} p^{y}(1-p)^{n-p}$.
Although the likelhood can be computed directly when replacing the
infinite limit with a finite value, we will demonstrate here that we
can easily evaluate it using a recursive algorithm that is both faster
and more numerical stable. The same idea is also applicable to the
negative Binomial case, and the case where we have replicated
observations of the same ``$n$''. We leave it to the reader to derive
these straight forward extentions.

The key observation is that both the Poisson and the Binomial
distribution can be evaluated recursively in $n$,
\begin{displaymath}
    \text{Pois}(n; \lambda) = \text{Pois}(n-1; \lambda) \frac{\lambda}{n}
\end{displaymath}
and
\begin{displaymath}
    \text{Binom}(y; n, p) = \text{Binom}(y; n-1, p) \frac{n}{n-y}(1-p),
\end{displaymath}
and then also for the Poisson-Binomial product
\begin{displaymath}
    \text{Pois}(n ; \lambda) \; \text{Binom}(y;  n, p)
    =
    \text{Pois}(n-1; \lambda) \; \text{Binom}(y; n-1, p)
    \frac{\lambda}{n-y}(1-p).
\end{displaymath}
If we define $f_i = \lambda(1-p)/i$ for $i=1, 2, \ldots$, we can make
use of this recursive form to express the likelihood with a finite
upper limit as
\begin{eqnarray}
    \text{Prob}(y) &=& \sum_{n = y}^{n_{\text{max}}}
                       \text{Pois}(n ; \lambda)\;
                      \text{Binom}(y;  n, p) \nonumber\\
                   &=& \text{Pois}(y; \lambda)\; \text{Binom}(y; y, p)
                       \Big\{ 1 + f_1 + f_1f_2 + 
                    \ldots
                      +f_1\cdots f_{n_\text{max}}
                       \Big\} \nonumber\\
                   &=& \text{Pois}(y; \lambda)\; \text{Binom}(y; y, p)
                       \Big\{ 1 + f_1(1+f_2(1+f_3(1+ \dots)))\Big\}\nonumber
\end{eqnarray}
The log-liklihood can then evaluated using the following simple
\texttt{R}-code
\begin{verbatim}
  fac = 1; ff = lambda * (1-p)
  for (i in (n.max - y):1) fac = 1 + fac * ff / i
  log.L = dpois(y, lambda, log=TRUE) +
          dbinom(y, y, p, log=TRUE) + log(fac)
\end{verbatim}
Since this evaluation is recursive in decreasing $n$, we have to chose
the upper limit $n_\text{max}$ up-front, for example as an integer
larger than $y$ so that $\frac{\lambda (1-p)}{n_\text{max}-y}$ is
small. Note that we are computing \texttt{fac} starting with the
smallest contributions, which are more numerical stable.

\end{document}

% LocalWords:

%%% Local Variables: 
%%% TeX-master: t
%%% End: